---
title: "Project 1"
author: "Sai Praneeth Vupputuri"
date: "25/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



Importing the data after setting the working directory and importing readr library

```{r}
library(readr)
library(dplyr)
setwd("~/Coursera/reproducible research/repdata_data_activity")
activity <- read_csv("activity.csv", na = "NA")
```
Converting date column into the appropriate format and aggregating the data date wise using the aggregate function
```{r}
activity$date <- as.Date(activity$date, "%Y-%m-%d")
final_data <- aggregate(steps~date, activity,sum)
```

Using the "final_data" dataframe to build a histogram with the base plot system

```{r, echo=TRUE}
hist(final_data$steps, xlab = "Number of total steps", ylab = "Number of Days", main = "Histogram of Total steps")
```

Displaying the mean  of the aggregated dataset
```{r}
aggregate(steps~date,activity,mean)
```

Displaying the  median of the aggregated dataset
```{r}
aggregate(steps~date,activity,FUN = "median")
```

Calculating the mean of steps for each interval 
```{r}
meaninterval <- aggregate(steps~interval,activity, mean)
```

Plotting the mean interval data
```{r}
plot(meaninterval, type = "l", xlab= "Intervals", ylab = "Steps", main = "Mean number of steps for 5 Second Intervals")
```

Finding which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max <- which.max(meaninterval$steps)
meaninterval[max,]
```
As you can see the maximum steps from average is 206 for interval 835

Finding the number of NA values using table function 
```{r}
table(is.na(activity$steps))
```
As shown abovw, the total number of NA values is 2304

Before we can substitute, I am first separting out the data into two parts, With NA values and without NA values
```{r}
clean <- activity[!is.na(activity$steps),]
nadata<- activity[is.na(activity$steps),]
```
Then merging the nadata dataframe and meaninterval dataframe
``` {r}
newdata<-merge(nadata, meaninterval, by=c("interval"))
names(newdata) = sub("steps.y", "steps", names(newdata))
newdata$steps.x<- NULL
```
Finally Merging clean dataframe with the mean-populated na dataset
```{r}
updatedset<- rbind(clean,newdata)
```
Checking if there are still any NA values in the Updated dataset
```{r}
table(is.na(updatedset$steps))
```
The lack of any true values in the table indicates that all missing NA values have been successfully replaced
```{r}
updatedset_final <- aggregate(steps~date, updatedset, sum)
hist(updatedset_final$steps, xlab = "Number of total steps", ylab = "Number of Days", main = "Histogram of Total steps")
```


Displaying the mean  of the new dataset
```{r}
aggregate(steps~date,updatedset_final,mean)
```

Displaying the  median of the new dataset
```{r}
aggregate(steps~date,updatedset_final,FUN = "median")
```
Creating a new "day" column and using that to create factor column that will contain weekend and weekday information
```{r}
updatedset$day <- weekdays(as.Date(updatedset$date))
updatedset$DayCategory <- ifelse(updatedset$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
head(updatedset)
```

```{r}
weekend <- filter(updatedset, updatedset$DayCategory == "Weekend")
weekend_final <- aggregate(steps~interval, weekend, mean)
weekday <- filter(updatedset, updatedset$DayCategory == "Weekday")
weekday_final <- aggregate(steps~interval, weekday, mean)
par(mfrow=c(2,1))
plot(weekday_final, type = "l", main = "weekday")
plot(weekend_final, type = "l", main = "weekend")
```

